---
title: "Using the acs .rds files"
author: "Chirag J Patel"
date: "12/1/2022"
output:
  html_document:
    df_print: paged
---


# How to use the Census ACS 5 year .rds files


```{r}
library(tidycensus)
library(tidyverse)
directory <- 'out'
variables <- load_variables(
  year=2020,
  dataset = c("acs5"),
  cache = FALSE
) 
variables <- variables %>% mutate(tablename=sapply(strsplit(variables$name, "\\_"), function(arr) {arr[[1]]})) # comprehensive listing of all variables available in the acs 5 year

```

# Stitch together the .rds files in a directory 

- See ehec_candidates.R  for more information 

```{r list rds files}
acs_files <- list.files(directory)
acs_table <- map(acs_files, ~read_rds(file=file.path(directory,.))) %>% bind_rows() # sf file
```

# Derived data 

- Derived data are rolled up to percentages (see file: ehec_candidates.R)
- begin with column character "d"

```{r derived data}
acs_table %>% select(starts_with("d")) %>% head()
```


```{r median income}
acs_table %>% ggplot(aes(d_median_income)) + geom_histogram() + facet_wrap(vars(state), nrow = 2)

acs_table %>% select(d_median_income, NAME) %>% filter(d_median_income > 200000)

```

# Simple correlations 
- cognitive impairment for individuals aged over 65 across all states vs. percent who are living under poverty of the older population

```{r}
acs_table %>% select(d_cognitive_female_over65_percentage, d_percent_older_under_poverty) %>% ggplot(aes(d_percent_older_under_poverty,d_cognitive_female_over65_percentage)) + geom_point(alpha=.1)

acs_table %>% select(d_unemployed_percentage, d_median_income) %>% ggplot(aes(d_unemployed_percentage,d_median_income)) + geom_point(alpha=.1)

```

# PLACES 2020

https://www.cdc.gov/places/index.html
Consists of small-area (census tract) estimates of prevalence


```{r places merge}
places <- read_csv("./PLACES__Local_Data_for_Better_Health__Census_Tract_Data_2022_release.csv.zip")
unique(places$Measure)

p_wide <- places %>% select(LocationName, MeasureId, Data_Value) %>% pivot_wider(names_from=MeasureId, values_from = Data_Value) # make in wide format, disease prevalence in each column

acs_table <- acs_table %>% left_join(p_wide, by=c("GEOID"="LocationName"))

acs_table %>% select(STROKE, d_cognitive_male_over65_percentage) %>% ggplot(aes(d_cognitive_male_over65_percentage,STROKE)) + geom_point()

acs_table %>% select(GHLTH, d_cognitive_male_over65_percentage) %>% ggplot(aes(d_cognitive_male_over65_percentage,GHLTH)) + geom_point()


```
